\documentclass[10pt,a4paper,final,oneside,openany,article]{memoir}

\input{prelude}

\title{Computational Geometry, Assignment 1}
\author{
    Jonas Brunsgaard \& Silas Ørting
}

\date{18th February 2015}

\bibliography{bibliography}

\begin{document}
\maketitle

\chapter*{Assignment 1.3}
\begin{enumerate}
  \item Start by sorting the points by $x$ coordinate.
  \item Add and edge $e$ between the two points with the most left
    $x$-coordinate and the most right $x$-coordinate.
  \item Create two empty lists $L_{\textrm{upper}}$ and $L_{\textrm{lower}}$.
  \item Iterate over the sorted list of points and determine whether each point
    lies above $e$. If the point is above, append the point to
    $L_{\textrm{upper}}$ otherwise append the point to $L_{\textrm{lower}}$.
  \item Reverse $L_{\textrm{lower}}$ and concatenate to $L_{\textrm{upper}}$
\end{enumerate}

\chapter*{Assignment 1.6}
\paragraph{a} The line segments will be a convex combination of the two
endpoints, and thus the line segments will also be in the set of all convex
combinations of all the endpoints, which is the definition of the convex hull.
$$ P \supseteq S \supseteq \mathcal{CH}_P $$
Also, we know that the points are on the lines and thus we can regard the
points as being a seubset of the lines, which again is a subset of the convex
hull to the lines.
$$ S \supseteq P \supseteq \mathcal{CH}_S $$
Thus,
$$ \mathcal{CH}_S = \mathcal{CH}_P $$
\paragraph{b} Todo 

\chapter*{Chan}
The algorithm combines the approach of Grahams scan and Jarvis’s march. The
idea is to partition the points into $m$ smaller sets of size $\rceil m/n
\lceil$. And then use Grahams scan to construct a convex hull for each of the
subsets. Afterwards the idea of Jarvis march is applied to wrap around the
convex hulls of the subsets. Because tangent finding between a point and a
convex polygon can be done in logarithmic time this approach is very efficient
and the result is a convex hull for all the points.


\chapter*{3.8}
When we insert a new edge in the DCEL we split a face into two faces, this means we have to update the incident-face information in $O(n)$ half-edge records. In the case of the polygon-partitioning algorithm, we do not need to store the face information, so adding a half-edge only requires that we make local updates to the four half-edges of the face we are splitting that has origin and destination at the vertices that are connected by the new edge. 

\chapter*{3.14 Visibility polygon}
Let $P$ be the polygon that is visible from $q$. If we traverse the edges of $P$ in CCW order then it holds that the angle between the lines $qp_i$ and $qp_{i+1}$ lie in $[0, \pi]$ and that angular sum is $2\pi$.

If we can find a point $p_0$ that is visible from $q$ then we can find the visibility polygon by starting in $p_0$ and adding vertices that satisfy the local angle constraint.

When a vertex $p_i$ that does not satisfy the local constraint is encountered we need to consider two cases. 
\begin{enumerate}
\item We have just turned right, then $p_i$ is obscured and we skip it.
\item We have just turned left, then the line to $p_i$ obscures some of the previous vertices, and we need to backtrack and discard previously accepted vertices untill the constraint is satisfied or we make a right turn. 
\end{enumerate}

The global angle constraint is eventually satisfied because we start from an initially visible vertex. If we have have violated the global angle constraint, then at some point we need to violate the local constraint in order to end up at the initally visible vertex.

\paragraph{}
Let {\verb visible } be a stack containing the vertices believed to be visible. Initially this contains the known visible point. Pseudocode is given below
\begin{verbatim}
p_1 = p_0
for i = 1 to N
  p = p_1.next
  skip = false
  while angle(q,p_1,q,p) > pi
    if is_left_turn(p_1.prev, p_1, p)
      p_1 = p_1.prev
      p.prev = p_1
    else
      p = p.next
      p_1.next = p
      p.prev = p_1
      skip = true
    end
  end
  if not skip
    p_1 = p
  end
end
\end{verbatim}
This runs in linear time, since each vertex is added at most once and removed at most once, and the angle constraint and turn direction can be computed in constant time.

\paragraph{}
Finding the visible starting point can be done in $O(n \log n)$ by triangulating the polygon and locating the triangle that $q$ falls in. We can then pick any of the three vertices as $p_0$.

\end{document}
